### 2024-04-19 docs/SOSP 2024/abstract.tex
pass

\begin{abstract}

  The rapid proliferation of machine learning together with the accelerating evolution of 
  the hardware ecosystem has led to a surge in the demand for model inference on a variety
  of hardware. This paper is motivated by the problems encountered when targeting 
  inference of decision tree based models, the most popular models on tabular data,
  to run at peak performance on diverse CPU and GPU targets. Existing solutions are neither
  portable nor achieve the best possible performance for the specific hardware they 
  target because they do not explore and customize optimization configurations to the model 
  being used.

  We present \Treebeard{}, a \emph{schedule-guided},
  \emph{retargetable} compiler for decision tree based models
  that searches over several optimization choices and automatically generates 
  high-performance inference routines for CPUs and GPUs.  
  \Treebeard{} has two core components. The first is a scheduling language 
  that encapsulates the optimization space, and techniques to efficiently
  explore this space. The second is an optimizing retargetable compiler that 
  can generate code for any specified schedule. \Treebeard{}'s ability to use different data layouts, 
  loop structures and caching strategies enables it to achieve portable performance
  across a range of CPUs and GPUs.   
  % To accomplish this level of configurability, we re-architect 
  % and significantly extend the open-source \TreebeardOLD{} CPU compiler
  % to support (i) schedule-guided compilation, (ii) retargetable GPU code
  % generation, and (iii) GPU-specific optimizations. 
  %The scheduler and retargetable code generator components work hand-in-glove, synergizing and mutually benefiting from each other.

>>>>

\begin{abstract}
  The proliferation of machine learning together with the rapid evolution of 
  the hardware ecosystem has led to a surge in the demand for model inference on a variety
  of hardware. 
  Decision tree based models are the most popular models on tabular data.
  This paper is motivated by the problems encountered when targeting 
  inference of these models to run at peak performance on CPU and GPU targets.
  Existing solutions are neither
  portable nor achieve the best possible performance for the specific hardware they 
  target. This is because they do not explore and customize optimization 
  configurations to the model being used.

  We present \Treebeard{}, a \emph{schedule-guided},
  \emph{retargetable} compiler for decision tree based models
  that searches over several optimization choices and automatically generates 
  high-performance inference routines for CPUs and GPUs.  
  \Treebeard{} has two core components. The first is a scheduling language 
  that encapsulates the optimization space, and techniques to efficiently
  explore this space. The second is an optimizing retargetable compiler that 
  can generate code for any specified schedule. \Treebeard{}'s ability to use different data layouts, 
  loop structures and caching strategies enables it to achieve portable performance
  across a range of targets.   
  % To accomplish this level of configurability, we re-architect 
  % and significantly extend the open-source \TreebeardOLD{} CPU compiler
  % to support (i) schedule-guided compilation, (ii) retargetable GPU code
  % generation, and (iii) GPU-specific optimizations. 
  %The scheduler and retargetable code generator components work hand-in-glove, synergizing and mutually benefiting from each other.

### 2024-04-19 docs/SOSP 2024/intro.tex
pass

%\TODO{cover Rapids, Tahoe, XGBoost and Treebeard}.  
Evaluation on a diverse set of models highlights that the best implementation often requires a careful composition 
of many optimization strategies like data layout optimizations, 
loop transformations, parallelization, and memory access optimizations. 
%For example, simultaneously traversing multiple trees with many threads, while managing the working set is 
%critical to get scalable performance. 
%This requires a combination of techniques like data layout optimizations, loop transformations, and memory access optimizations. 
%Existing systems each use specific data layouts for the underlying tree and use fixed loop structures. 
%Parallelism, loop structure and model representation together determine the accesses patterns and the working set. 
Existing systems today are mostly library based, and only support a predefined combination of optimizations
and typically only target a single platform. 
On the CPU, XGBoost~\cite{XGBoost} uses a sparse representation for the model 
and a loop structure that processes one tree for a block of rows before moving to 
the next tree. On the GPU, RAPIDS FIL~\cite{FIL} uses a reorg representation~\cite{FIL}
and partitions trees across a fixed number of threads. 
%% RG: 
% \TODO{Should we say RAPIDS is only for nVidia GPUs?} 
%%
Tahoe~\cite{Tahoe} uses a variation of the reorg representation and 
has four predefined inference strategies from which it picks one based on an analytical model.
The GPU systems are CUDA based and only work on NVIDIA GPUs.
%These systems are all library based and are hard to extend. 
\TreebeardOLD{}, the state-of-the-art decision tree model compiler for CPUs 
%% RG
built using the MLIR infrastructure~\cite{MLIR}, supports two fixed loop structures 
% and does not scale well with increasing number of threads. Additionally, it 
and lacks GPU specific optimizations that are critical to scale performance to massive number of threads.
% \TODO{Shouldn't we reverse this? First say no GPU support and then the scaling issue.}

>>>>

%\TODO{cover Rapids, Tahoe, XGBoost and Treebeard}.  
Evaluation on a diverse set of models highlights that the best implementation often requires a careful composition 
of many optimization strategies like data layout optimizations, 
loop transformations, parallelization, and memory access optimizations. 
%For example, simultaneously traversing multiple trees with many threads, while managing the working set is 
%critical to get scalable performance. 
%This requires a combination of techniques like data layout optimizations, loop transformations, and memory access optimizations. 
%Existing systems each use specific data layouts for the underlying tree and use fixed loop structures. 
%Parallelism, loop structure and model representation together determine the accesses patterns and the working set. 
Existing systems today are mostly library based, only support a predefined combination of optimizations
and typically only target a single platform. 
On the CPU, XGBoost~\cite{XGBoost} uses a sparse representation for the model 
and a loop structure that processes one tree for a block of rows before moving to 
the next tree. On the GPU, RAPIDS FIL~\cite{FIL} uses a reorg representation~\cite{FIL}
and partitions trees across a fixed number of threads. 
%% RG: 
% \TODO{Should we say RAPIDS is only for nVidia GPUs?} 
%%
Tahoe~\cite{Tahoe} uses a variation of the reorg representation and 
has four predefined inference strategies from which it picks one based on an analytical model.
The GPU systems are CUDA based and only work on NVIDIA GPUs.
%These systems are all library based and are hard to extend. 
\TreebeardOLD{}, the state-of-the-art decision tree model compiler for CPUs
%% RG
, supports two fixed loop structures 
% and does not scale well with increasing number of threads. Additionally, it 
and lacks GPU specific optimizations that are critical to scale performance to massive number of threads.
% \TODO{Shouldn't we reverse this? First say no GPU support and then the scaling issue.}
